On exponential sums arising from the classical doubling method

### 2025-04-02 main.tex
And I forgot to remove \epsilon_H
Explaining the signs

\subsubsection{Split torus computation II}
We use the results from the previous section to compute the doubling method Jacobi sum for a torus of the form $\multiplicativegroup{\quadraticFieldExtension{k}}$. Let $\alpha \colon \multiplicativegroup{\quadraticFieldExtension{k}} \to \multiplicativegroup{\cComplex}$ and $\chi \colon \multiplicativegroup{\quadraticExtension} \to \multiplicativegroup{\cComplex}$ be characters such that $\involutionPlusOne{\chi} \ne 1$. Our goal is to compute $$\sum_{\substack{x \in \multiplicativegroup{\quadraticFieldExtension{k}}\\
		x \ne -1}} \alpha \left(x\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + x\right)\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + \minusInvolution{x}\right)\right).$$
This can be rewritten as
$$\sum_{\substack{x \in \multiplicativegroup{\quadraticFieldExtension{k}}\\
		x \ne -1}} \alpha \left(x\right) \minusInvolution{\chi}\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(x\right)\right) \involutionPlusOne{\chi}\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + x\right)\right).$$
By the previous section, this equals
\begin{equation*}
	-\sizeof{\quadraticExtension}^{\frac{k}{2}} \GaussSumCharacter{\alpha}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\alpha^{-1}}{\involution{\chi}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumSingleCharacter{\minusInvolutionMinusOne{\chi}}{\fieldCharacter_{\quadraticExtension}}^k,
\end{equation*}
which equals
\begin{equation*}
	-\sizeof{\quadraticExtension}^{\frac{k}{2}} \GaussSumCharacter{\involution{\alpha}}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\minusInvolution{\alpha}}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumSingleCharacter{\minusInvolutionMinusOne{\chi}}{\fieldCharacter_{\quadraticExtension}}^k.
\end{equation*}
Thus we obtained the following.
\begin{theorem}We have the identity
\begin{align*}
	& \sum_{\substack{x \in \multiplicativegroup{\quadraticFieldExtension{k}}\\
			x \ne -1}} \alpha \left(x\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + x\right)\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + \minusInvolution{x}\right)\right) \\
	=& -\sizeof{\quadraticExtension}^{\frac{k}{2}} \GaussSumCharacter{\alpha}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\minusInvolution{\alpha}}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumSingleCharacter{\minusInvolutionMinusOne{\chi}}{\fieldCharacter_{\quadraticExtension}}^k.
\end{align*}
\end{theorem}

@@

\begin{theorem}\label{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}
	Suppose that $T = T_G \cong \prod_{j=1}^r \multiplicativegroup{\quadraticFieldExtension{k_j}} \times \prod_{i=1}^s \finiteFieldExtension{2m_i}^1$ is a maximal torus of $G$ and let $T_H$ be the unique maximal torus of $H$ such that $T_H \cap G = T_G$. Suppose that $\theta \colon T_G \to \multiplicativegroup{\cComplex}$ is a character and that under the above isomorphism $\theta = \alpha_1 \times \dots \times \alpha_r \times \theta_1 \times \dots \times \theta_s$, where $\alpha_j \colon \multiplicativegroup{\quadraticFieldExtension{k_j}} \to \multiplicativegroup{\cComplex}$ and $\theta_i \colon \finiteFieldExtension{2m_i}^1 \to \multiplicativegroup{\cComplex}$ are characters for every $i$ and $j$. Denote \begin{align*}
		g_T\left(\chi, \theta, \fieldCharacter\right)
		= & \prod_{i=1}^s \GaussSumCharacter{\theta'_i}{\chi}{\fieldCharacter_{2m_i}} \cdot \prod_{j=1}^r \GaussSumCharacter{\minusInvolution{\alpha_j}}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k_j}}} \GaussSumCharacter{\alpha_j}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k_j}}}.
	\end{align*} Then
	\begin{align*}
		 \dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \varepsilon_{\algebraicGroup{H}} R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_T\left(\chi, \theta, \fieldCharacter\right),
	\end{align*}
	and \begin{align*}
		\dblNegVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \begin{dcases}
			0 & \dim_{\finiteField} \hermitianSpace \text{ is odd.}\\
			\dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} \cdot \prod_{i=1}^s \theta_i\left(-1\right) \cdot \prod_{j=1}^r \alpha_j\left(-1\right) & \text{ otherwise.}
		\end{dcases}
	\end{align*}
\end{theorem}
\begin{proof}
	Notice that if $\dim_{\finiteField} \hermitianSpace$ is odd then $\negJacobiKernel{\chi}\left(t\right) = 0$ for any $t \in T_{G}$, as $\idmap_{\hermitianSpace} - t$ always has determinant zero.

@@

	Hence our theorem follows from the results mentioned above and from the formula $$R_T\left(1\right) = \varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{H}}} \frac{\grpIndex{H}{T}}{\sizeof{H}_p},$$ the fact that $$ \sizeof{\lieAlgebra} = \sizeof{H}_p^2 \cdot q^{\left\lfloor\frac{\dim_{\finiteField} \hermitianSpace}{2}\right\rfloor},$$
	the fact that \begin{equation*}
		\varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{H}}} = \varepsilon_{\algebraicGroup{G}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}},
	\end{equation*}
	and the facts
	\begin{align*}
		\varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\left\lfloor \frac{1}{2}\dim_{\finiteField} \hermitianSpace \right\rfloor} & \,\,\,\,\,\text{and}\,\,\,\,\, \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}} = \left(-1\right)^r.
	\end{align*}
\end{proof}
\begin{remark}
	We have $$\varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\left\lfloor \frac{\dim_{\finiteField} \hermitianSpace}{2}\right\rfloor}.$$ If $\finiteField = \quadraticExtension$ and $\dim_{\finiteField} \hermitianSpace$ is even then we also have $$\varepsilon_{\algebraicGroup{\GroupExtension{G}}} = \left(-1\right)^{\frac{\dim_{\finiteField} \hermitianSpace}{2} + 1}.$$
\end{remark}
\begin{remark}
	Suppose that $\pi$ is any irreducible representation of $H$ such that $\trace \pi$ appears with non-zero coefficient in $R_{T,\theta}$. Then the central character of $\pi$ satisfies
	\begin{align*}
		\centralCharacter{\pi}\left(-1\right) = \theta\left(-1\right) =& \prod_{j=1}^r \alpha_j\left(-1\right) \cdot \prod_{i=1}^s \theta_i\left(-1\right).
	\end{align*}
	Thus \begin{align*}
		\dblNegVirtualJacobiSumScalar{R_{T, \theta}}{\chi} =& R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) \centralCharacter{\pi}\left(-1\right) g_T\left(\chi, \theta, \fieldCharacter\right).
	\end{align*}
\end{remark}

>>>>

\subsubsection{Split torus computation II}
We use the results from the previous section to compute the doubling method Jacobi sum for a torus of the form $\multiplicativegroup{\quadraticFieldExtension{k}}$. Let $\alpha \colon \multiplicativegroup{\quadraticFieldExtension{k}} \to \multiplicativegroup{\cComplex}$ and $\chi \colon \multiplicativegroup{\quadraticExtension} \to \multiplicativegroup{\cComplex}$ be characters such that $\involutionPlusOne{\chi} \ne 1$. Our goal is to compute $$\sum_{\substack{x \in \multiplicativegroup{\quadraticFieldExtension{k}}\\
		x \ne -1}} \alpha \left(x\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + x\right)\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + \minusInvolution{x}\right)\right).$$
This can be rewritten as
$$\sum_{\substack{x \in \multiplicativegroup{\quadraticFieldExtension{k}}\\
		x \ne -1}} \alpha \left(x\right) \minusInvolution{\chi}\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(x\right)\right) \involutionPlusOne{\chi}\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + x\right)\right).$$
By the previous section, this equals
\begin{equation*}
	-\sizeof{\quadraticExtension}^{\frac{k}{2}} \GaussSumCharacter{\alpha}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\alpha^{-1}}{\involution{\chi}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumSingleCharacter{\minusInvolutionMinusOne{\chi}}{\fieldCharacter_{\quadraticExtension}}^k,
\end{equation*}
which equals
\begin{equation*}
	-\sizeof{\quadraticExtension}^{\frac{k}{2}} \GaussSumCharacter{\involution{\alpha}}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\minusInvolution{\alpha}}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumSingleCharacter{\minusInvolutionMinusOne{\chi}}{\fieldCharacter_{\quadraticExtension}}^k.
\end{equation*}
Thus we obtained the following.
\begin{theorem}\label{thm:split-case-deligne-lusztig-computation}We have the identity
\begin{align*}
	& \sum_{\substack{x \in \multiplicativegroup{\quadraticFieldExtension{k}}\\
			x \ne -1}} \alpha \left(x\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + x\right)\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k} \slash \quadraticExtension}\left(1 + \minusInvolution{x}\right)\right) \\
	=& -\sizeof{\quadraticExtension}^{\frac{k}{2}} \GaussSumCharacter{\alpha}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\minusInvolution{\alpha}}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumSingleCharacter{\minusInvolutionMinusOne{\chi}}{\fieldCharacter_{\quadraticExtension}}^k.
\end{align*}
\end{theorem}

@@

\begin{theorem}\label{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}
	Suppose that $T = T_G \cong \prod_{j=1}^r \multiplicativegroup{\quadraticFieldExtension{k_j}} \times \prod_{i=1}^s \finiteFieldExtension{2m_i}^1$ is a maximal torus of $G$ and let $T_H$ be the unique maximal torus of $H$ such that $T_H \cap G = T_G$. Suppose that $\theta \colon T_G \to \multiplicativegroup{\cComplex}$ is a character and that under the above isomorphism $\theta = \alpha_1 \times \dots \times \alpha_r \times \theta_1 \times \dots \times \theta_s$, where $\alpha_j \colon \multiplicativegroup{\quadraticFieldExtension{k_j}} \to \multiplicativegroup{\cComplex}$ and $\theta_i \colon \finiteFieldExtension{2m_i}^1 \to \multiplicativegroup{\cComplex}$ are characters for every $i$ and $j$. Denote \begin{align*}
		g_T\left(\chi, \theta, \fieldCharacter\right)
		= & \prod_{i=1}^s \GaussSumCharacter{\theta'_i}{\chi}{\fieldCharacter_{2m_i}} \cdot \prod_{j=1}^r \GaussSumCharacter{\minusInvolution{\alpha_j}}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k_j}}} \GaussSumCharacter{\alpha_j}{\chi}{\fieldCharacter_{\quadraticFieldExtension{k_j}}}.
	\end{align*} Then
	\begin{align*}
		 \dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_T\left(\chi, \theta, \fieldCharacter\right),
	\end{align*}
	and \begin{align*}
		\dblNegVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \begin{dcases}
			0 & \dim_{\finiteField} \hermitianSpace \text{ is odd.}\\
			\dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} \cdot \prod_{i=1}^s \theta_i\left(-1\right) \cdot \prod_{j=1}^r \alpha_j\left(-1\right) & \text{ otherwise.}
		\end{dcases}
	\end{align*}
\end{theorem}
\begin{proof}
	Notice that if $\dim_{\finiteField} \hermitianSpace$ is odd then $\negJacobiKernel{\chi}\left(t\right) = 0$ for any $t \in T_{G}$, as $\idmap_{\hermitianSpace} - t$ always has determinant zero.

@@

	Hence our theorem follows from the results mentioned above and from the formula $$R_T\left(1\right) = \varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{H}}} \frac{\grpIndex{H}{T}}{\sizeof{H}_p},$$ the fact that $$ \sizeof{\lieAlgebra} = \sizeof{H}_p^2 \cdot q^{\left\lfloor\frac{\dim_{\finiteField} \hermitianSpace}{2}\right\rfloor},$$
	the fact that \begin{equation*}
		\varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{H}}} = \varepsilon_{\algebraicGroup{G}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}},
	\end{equation*}
	and the facts
	\begin{align*}
		\varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\left\lfloor \frac{1}{2}\dim_{\finiteField} \hermitianSpace \right\rfloor} & \,\,\,\,\,\text{and}\,\,\,\,\, \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}} = \left(-1\right)^r.
	\end{align*}
	In this computation, the factor $\varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}}$ comes from \Cref{thm:split-case-deligne-lusztig-computation} while the factor $\varepsilon_{\algebraicGroup{G}}$ comes from the definition of the normalization factor $c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)$.
\end{proof}
\begin{remark}
	We have $$\varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\left\lfloor \frac{\dim_{\finiteField} \hermitianSpace}{2}\right\rfloor}.$$ If $\finiteField = \quadraticExtension$ and $\dim_{\finiteField} \hermitianSpace$ is even then we also have $$\varepsilon_{\algebraicGroup{\GroupExtension{G}}} = \left(-1\right)^{\frac{\dim_{\finiteField} \hermitianSpace}{2} + 1}.$$
\end{remark}
\begin{remark}
	Suppose that $\pi$ is any irreducible representation of $H$ such that $\trace \pi$ appears with non-zero coefficient in $R_{T,\theta}$. Then the central character of $\pi$ satisfies
	\begin{align*}
		\centralCharacter{\pi}\left(-1\right) = \theta\left(-1\right) =& \prod_{j=1}^r \alpha_j\left(-1\right) \cdot \prod_{i=1}^s \theta_i\left(-1\right).
	\end{align*}
	Thus \begin{align*}
		\dblNegVirtualJacobiSumScalar{R_{T, \theta}}{\chi} =& R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) \centralCharacter{\pi}\left(-1\right) g_T\left(\chi, \theta, \fieldCharacter\right).
	\end{align*}
\end{remark}
